---
title: "Evaluating Oppurtunity and Risk for Lending Loans"
output: github_document
---

## Data Preparation & Cleaning

```{r setup, include=F}
knitr::opts_chunk$set(
  fig.path = "markdown_figs/data_preparation-",
  eval = F
)
```

```{r library, message=F}
library(magrittr)
library(dplyr)
library(data.table) # Faster reading speeds for csv files (fread)
```

## Computing Probability of Default for Test Set

### Loading the Test Datasets

```{r}
test_scenario3 <- fread("data/test_scenario3.csv", data.table = F, colClasses = c("character"))
test_scenario1_2 <- fread("data/test_scenario1_2.csv", data.table = F, colClasses = c("character"))
```

#### Getting Columns Specific to Scenario 1 & 2

```{r}
request_loan_sc1_2 <- test_scenario1_2 %>% select(requested_loan)
```

#### Getting Columns Specific to Scenario 3

```{r}
request_loan_sc3 <- test_scenario3 %>% select(requested_loan)
proposed_intrest_rate <- test_scenario3 %>% select(Proposed_Intrest_Rate)
```

#### Recoding The Variables as Numeric

```{r}
request_loan_sc1_2 <- request_loan_sc1_2 %>% mutate_all(as.numeric)
request_loan_sc3 <- request_loan_sc3 %>% mutate_all(as.numeric)
proposed_intrest_rate <- proposed_intrest_rate %>% mutate_all(as.numeric)
```

#### Loading PD and LGD of Customers in Test Set

```{r}
customer_LGD <- read.csv("data/customer-LGD.csv")
customer_PD <- read.csv("data/customer-PD.csv")
```


### Evaluating Customers in Scenario 1

CAPITAL <- 1.4e+09

```{r}
scenario1 <- test_scenario1_2 %>%
  select(id) %>%
  bind_cols(request_loan_sc1_2,customer_PD, customer_LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Oppurtunity = requested_loan * 0.0432 * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD
  )

summary(scenario1)
```

#### Total Requested Loan in Scenario 1

```{r}
scenario1 %>%
  select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

#### Total Sum of Oppurtunity w/ a Expected Default Rate in Scenario 1

##### PD at 0.5 in Scenario 1

```{r}
scenario1 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  select(default) %>%
  summary()
```

```{r}
scenario1 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

##### PD at 0.1 in Scenario 1

```{r}
scenario1 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  select(default) %>%
  summary()
```

```{r}
scenario1 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

#### Approving Customers in Scenario 1 based on Oppurtunity >= Risk

```{r}
scenario1 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  select(loan_approval) %>%
  summary()
```

##### Total Sum of Oppurtunity in Scenario 1

```{r}
scenario1 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

##### Total Requested Loan where Oppurtunity >= Risk in Scenario 1

```{r}
scenario1 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

### Final Descision on Customer Loan Approval in Scenario 1

```{r}
scenario1 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  select(loan_approval) %>%
  write.csv(file = "data/G2_S1.csv", row.names = F)
```

### Evaluating Customers in Scenario 2

CAPITAL <- 450e+6

```{r}
scenario2 <- test_scenario1_2 %>%
  select(id) %>%
  bind_cols(request_loan_sc1_2,customer_PD, customer_LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Oppurtunity = requested_loan * 0.0432 * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD
  )
```

#### Final Descision on Customer Loan Approval in Scenario 2

```{r}
scenario2 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  select(loan_approval) %>%
  write.csv(file = "data/G2_S2.csv", row.names = F)
```

### Evaluating Customers in Scenario 3

CAPITAL <- 1.4e+09

```{r}
scenario3 <- test_scenario3 %>%
  select(id) %>%
  bind_cols(request_loan_sc1_2, proposed_intrest_rate, customer_PD, customer_LGD) %>%
  rename(LGD = x) %>%
  mutate(
    Oppurtunity = requested_loan * Proposed_Intrest_Rate * 5 * (1 - PD),
    Risk = requested_loan * PD *  LGD
  )

head(scenario3)
```

```{r}
scenario3 %>% select(Proposed_Intrest_Rate) %>% summary()
```

### Total Requested Loan where Oppurtunity >= Risk in Scenario 3

```{r}
scenario3 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  select(requested_loan) %>%
  transmute(total_requested_loan = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

#### Total Sum of Oppurtunity in Scenario 3

```{r}
scenario3 %>%
  mutate(loan_approval = ifelse(Oppurtunity >= Risk, 1, 0)) %>%
  mutate_at(c("loan_approval"), as.factor) %>%
  filter(loan_approval == 1) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

#### Total Sum of Oppurtunity w/ a Expected Default Rate in Scenario 3

##### PD at 0.5 in Scenario 3

```{r}
scenario3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  select(default) %>%
  summary()
```

```{r}
scenario3 %>%
  mutate(default = ifelse(PD >= 0.5, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

##### PD at 0.1 in Scenario 3

```{r}
scenario3 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  select(default) %>%
  summary()
```

```{r}
scenario3 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate_at(c("default"), as.factor) %>%
  filter(default == 0) %>%
  select(Oppurtunity) %>%
  transmute(total_oppurtunity = format(sum(.), big.mark = ",", scientific = F)) %>%
  head(1)
```

### Final Descision on Customer Loan Approval in Scenario 3

```{r}
scenario3 %>%
  mutate(default = ifelse(PD >= 0.1, 1, 0)) %>%
  mutate(loan_approval = ifelse(default == 0, 1, 0)) %>%
  select(loan_approval) %>%
  write.csv(file = "data/G2_S3.csv", row.names = F)
```
